A Survey Wearable Device Risk Perceptions

### 2015-01-05 wearables_survey.tex
Minor edits in beginning of analysis

\subsection{Ubiquitous Sensing}
At an unprecedented rate, we are moving towards automated capture and ubiquitous access~\cite{abowd2000charting}, whether it be through smartphones, smarthomes, networked devices, or, more recently, wearables. Wearbles have new capabilities which allow capture of unique data about the user and his/her surroundings. As information capture becomes pervasive and subtle, people are naturally concerned about privacy and security concerns ~\cite{palen2003unpacking}. Privacy in ubiquitous systems can be achieved my adhering to principles such as notice, choice, and consent ~\cite{langheinrich2001privacy}.

\subsection{Smartphones and Wearables Concerns}
There exists work which addresses anticipated privacy and security challenges related to mobile and wearable devices from a system standpoint, such as \cite{DiPietro}. However, works like these are not grounded in user research and do not make claims about what the the most relevant concerns for the average user are. Felt et. al's study of smartphone user concerns \cite{Felt} was the first to investigate the smartphone security and privacy landscape from a user study standpoint. The survey asked 3,115 smartphone users about 99 risks associated with various smartphone privileges. Participants were asked how upset they would be if a certain action had occurred without their permission. Participants rated each situation on a likert scale ranging from indifferent (1) to very upset (5). Our work is motivated by this mobile study, and aims to also investigate wearables security and privacy with a user focus to find the most relevant challenges. We use a similar scale to ask our users how upset they would be if a certain action was taken without their permission, but ask questions about new capabilities on wearable devices. To our knowledge, our work is the first of its kind to investigate such user concerns.

\subsection{Technology Risk Perceptions}
While evaluating which are the most pressing concerns for users, it is also important to take into account how users' perception of particular technology. A seminal risk-benefit analysis psychometric study by Fischoff \cite{Fischhoff} studied how 76 users perceived the risks and benefits of 30 activities and technologies. In the study, participants were prompted to rate the risks and benefits of 30 activities and technologies for all people affected by the technology, for a long period of time. The participants rated these technologies with respect to each other on a numerical scale, being instructed to rate the least risky or least beneficial technology a 10 and scaling the ratings linearly (i.e. a technology with risk rating 20 would be considered twice as risky compared to a technology with a risk rating of 10). In our survey, we use a similar prompt to ask participants to rate the risks and benefits of 20 new technologies. Denning et. al conducted a study investigating privacy perspectives the bystanders of wearable devices ~\cite{Denning2014}. Bystanders expressed concerns due to the subtle, easy-to-use, and prevalent nature of wearables. While people have conducted studies on privacy concerns with respect to wearables, we are aware of no work which as looked at privacy perspectives of the owner of the device. In our survey, we show how risk and benefit perception plays a role in how users perceive various new devices and new technologies in wearables.

@@

In performing our analysis in the next section, we chose to focus on the very upset rate (VUR) of various scenarios (i.e., the percentage of participants who reported a `5' on the Likert scales), rather than the average of all Likert scores. We did this for the same reasons as Felt {\it et al.}: the VUR does not presume that the ratings, ranging from ``indifferent'' to ``very upset,'' are linearly spaced. Additionally, we believe that everyone would be upset, at least a little, at the given scenarios (i.e., in all scenarios, the device takes action without permission). Thus, we believe that the main distinguishing factor of a participant reacting to a given scenario is if they were maximally upset or not, rather than how upset they were. In the risk/benefit section, followed Fischhoff {\it et al.}'s methodology and did not normalize the numerical responses. Rather, we report medians and quartiles, which are not heavily impacted by outliers. Finally, for the open-ended question at the end, two researchers independently coded 1,785 responses and had an initial agreement rate of 89.78\%. The researchers discussed any disagreements, and resolved them so that the final codings reflect unanimous agreement.

@@

In this section, we present the results of our multiple-part survey. We first discuss participants responses to the various data-sharing scenarios, and how data type, device type, and recipient contributed to how threatening a situation was perceived by participants. Next, we discuss participants' risk/benefit assessment of various new technologies with respect to well-established technologies. We conclude the section with participants' self-reported concerns about the biggest risks in owning wearable devices.

>>>>

\subsection{Ubiquitous Sensing}
{\color{red} This is plagiarized from our CHI paper! REWRITE!!! } Wearable devices, along with many other devices, are moving towards automated capture and ubiquitous  access~\cite{abowd2000charting}. More than ever, people will interact with sensors while carrying out daily activities, from a myriad of sources and for a variety of reasons: ``lifelogging'' devices to record their daily lives~\cite{hoyle2014privacy,dickie2004augmenting}, memory augmentation devices~\cite{chen2010augmenting,devaul2003memory,kalnikaite2010now}, smart homes that optimize living conditions~\cite{chan2008review,chan2009smart,edwards2001home}, devices that create augmented reality environments~\cite{azuma1997survey,azuma2001recent}, and even devices integrated into clothing \cite{anliker2001weararm}. The ubiquitous capture and storage of information naturally raises concerns about the preservation of privacy~\cite{palen2003unpacking}. Privacy in the field of ubiquitous computing relies on principles including notice, choice, and consent~\cite{langheinrich2001privacy}.

\subsection{Smartphones and Wearables Concerns}
{\color{red} Add way more details about Felt's study, so that it's clear how we borrowed the methodology, without the reader having to read that paper.}
There exists work which addresses anticipated privacy and security challenges related to mobile and wearable devices from a system standpoint, such as \cite{DiPietro}. However, works like these are not grounded in user research to make claims to the most relevant concerns for the average user. Felt et. al's study of smartphone user concerns \cite{Felt} was the first to investigate the security and privacy research landscape from a user study standpoint. The survey asked 3,115 smartphone users about 99 risks associated with various smartphone privileges. Our work is motivated by this mobile study, and aims to also investigate wearables security and privacy with a user focus to find the most relevant challenges. To our knowledge, our work is the first of its kind to investigate user concerns for wearable devices.

\subsection{Technology Risk Perceptions}
{\color{red} Again, you need to add lots more about this paper, so that it's clear from your description what they did, and how we build upon it.}
While evaluating which are the most pressing concerns for users, it is also important to take into account how users' perception of particular technology. A seminal risk-benefit analysis psychometric study by Fischoff \cite{Fischhoff} studied how 76 users perceived the risks and benefits of 30 activities and technologies. In our survey, we also explore how risk and benefit perception plays a role in how users perceive various new devices and new technologies in wearables.

{\color{red} You really need to cite ~\cite{Denning2014} in this section.}

{\color{red} To make it flow, add 1-3 sentences here about the gaps in the related work, which then motivates our study (e.g., while people have looked at X, we are aware of no work that has looked at Y. Thus, we performed a survey to examine Y).}

@@

%SE: Are the above numbers correct? Also, can you add a range of ages?

In performing our analysis in the next section, we chose to focus on the very upset rate (VUR) of various scenarios (i.e., the percentage of participants who reported a `5' on the Likert scales), rather than the average of all Likert scores. We did this for the same reasons as Felt {\it et al.}: the VUR does not presume that the ratings, ranging from ``indifferent'' to ``very upset,'' are linearly spaced. Additionally, we believe that everyone would be upset, at least a little, at the given scenarios (i.e., in all scenarios, the device takes action without permission). Thus, we believe that the main distinguishing factor of a participant reacting to a given scenario is if they were maximally upset or not, rather than how upset they were. In the risk/benefit section, we followed Fischhoff {\it et al.}'s methodology and did not normalize the numerical responses. Rather, we report medians and quartiles, which are not heavily impacted by outliers. Finally, for the open-ended question at the end (i.e., additional privacy concerns), two researchers independently coded 1,785 responses and had an initial agreement rate of 89.8\%. The researchers discussed any disagreements, and resolved them so that the final codings reflect unanimous agreement.

@@

In this section, we present our survey results. We first discuss participants' responses to the various data-sharing scenarios, and how data type, device type, and recipient contributed to how threatening a situation was perceived. Next, we discuss participants' risk/benefit assessment of various new technologies relative to well-established technologies. We conclude the section with participants' self-reported concerns about the biggest risks in owning wearable devices.
